---
title: "SI model of a vaccine trial"
author: ""
date: '`r Sys.Date()`'
output: html_document
---

```{r setup, include=FALSE}
library("deSolve")
library("tidyverse")
library("EpiModel")
```

We are modeling a closed population SI model meant to simulate a vaccine trial. We are not modeling infections from I to S but rather only infections from the outside population, with infection rate based on the population prevalence `prev` (of viremic individuals), the exposure rate (serodiscordant sexual contacts per time) `c`, and the transmission rate (per contact) `p`. The per contact effect of vaccination is `epsilon`, and at this iteration of the model `epsilon` is not time-varying (the per contact vaccine effect does not decay over time). 

p = transmission rate (per contact)   
c = exposure rate (serodiscordant sexual contacts per time)  
prev = prevalence  (this should probably be prevalence of viremic individuals)  
lambda = p * c * prev  
epsilon = per contact vaccine efficacy; vaccine-induced reduction in the risk of HIV infection from a single exposure  

## Model setup

Basic functions:  

`dS/dt = -lambda*S`   
`dI/dt = lambda*S`  

Sp = susceptible placebo  
Ip = infected placebo  
Sv = susceptible vaccinated  
Iv = infected vaccinated  

Svh = susceptible vaccinated high exposure  
SvL = susceptible vaccinated low exposure  
Ivh = infected vaccinated high exposure  
Ivl = infected vaccinated low exposure  

```{r parameters}

p <- 0.01   #transmission rate (per contact)
c <- 0.25   #contact rate (contacts per time)
prev <- 0.20
lambda <- p*c*prev
epsilon <- 0.50 #per act vaccine efficacy
risk <- 3.0 #risk multiplier
```
  
  
```{r ODE model}

si_ode <- function(times, init, param){
  with(as.list(c(init, param)), {
    
    # Flows
    # the number of people moving from S to I at each time step
    SIp.flow <- lambda*Sp
    SIv.flow <- lambda*epsilon*Sv
    SIvh.flow <- risk*lambda*(1-epsilon)*Svh
    SIvl.flow <- lambda*(1-epsilon)*Svl
    
    # ODEs
    #placebo; homogeneous risk
    dSp <- -SIp.flow
    dIp <- SIp.flow  #lambda*Sp
    
    # vaccine; homogeneous risk
    dSv <- -SIv.flow
    dIv <- SIv.flow  #lambda*epsilon*Sv

    # vaccine; heterogeneous risk
    dSvh <- -SIvh.flow
    dIvh <- SIvh.flow  #risk*lambda*(1-epsilon)*Svh
    dSvl <- -SIvl.flow
    dIvl <- SIvl.flow  #lambda*(1-epsilon)*Svl

    #Output
    list(c(dSp,dIp,
           dSv,dIv,
           dSvh,dIvh,
           dSvl,dIvl,
           SIp.flow,SIv.flow,SIvh.flow,SIvl.flow))
  })
}
```
  
  
```{r EpiModel}

param <- param.dcm(lambda = lambda, epsilon = epsilon, risk = risk)
init <- init.dcm(Sp = 5000, Ip = 1,
                 Sv = 5000, Iv = 1,
                 Svh = 2500, Ivh = 1, 
                 Svl = 2500, Ivl = 1,
                 SIp.flow = 0, SIv.flow = 0, SIvh.flow = 0, SIvl.flow = 0)

control <- control.dcm(nsteps = 365*3, new.mod = si_ode)

mod <- dcm(param, init, control)
mod

df <- as.data.frame(mod)
head(df, 25)
```
  

```{r Incidence calculation}

df$years <- df$time/365  #Changing to years
df$Svh.Svl <- df$Svh + df$Svl   #Summing all susceptible in heterogeneous risk pop
df$Ivh.Ivl <- df$Ivh + df$Ivl   ##Summing all infected in heterogeneous risk pop


```

   
```{r plots}

par(mar = c(3,3,2,1), mgp = c(2,1,0))
plot(mod, y = c("Iv", "Ivh", "Ip", "Ivl"), 
     alpha = 0.8, 
     main = "Cumulative infections",
     legend = "full")
plot(mod, y = c("SIv.flow", "SIvh.flow", "SIp.flow", "SIvl.flow"), 
     alpha = 0.8, 
     main = "Daily infections",
     legend = "full")
```


