---
title: "SI model of an HIV vaccine trial"
author: ""
date: '`r Sys.Date()`'
output: html_document
---

```{r setup, include=FALSE}
library(deSolve)
library(tidyverse)
library(EpiModel)
library(survival)
```

Using models to assess the impact of vaccine leakiness on trial vaccine efficacy measures.

We are modeling a closed population SI deterministic compartmental model meant to simulate a vaccine trial. We are not modeling infections from I to S but rather only infections from the outside population, with infection rate based on the population prevalence `prev` (of viremic individuals), the exposure rate (serodiscordant sexual contacts per time) `c`, and the transmission rate (per contact) `p`. The per contact effect of vaccination is `epsilon`, and with this iteration of the model `epsilon` is: 1) not time-varying (the per contact vaccine effect does not decay over time); and 2) assumes a homogeneous effect (does not vary by mark / viral genotype). This model structure also removes the possibility of indirect effects from vaccination.

To do:  in the context of the AMP Trial and the different results seen in the sub-studies (703 vs 704); is this due to different forces of infection between the populations?

`beta` = transmission rate (per contact)   
`c` = exposure rate (serodiscordant sexual contacts per time)  
`prev` = prevalence  (prevalence of viremic individuals)  
`lambda = beta * c * prev`  
`epsilon` = per contact vaccine efficacy; vaccine-induced reduction in the risk of HIV infection from a single exposure  

We use an initial set of transmission parameters for South Africa borrowing from Alain Vandormael (2018):    

     "We used realistic parameter values for the SIR model, based on earlier HIV studies that have been undertaken in the sub-Saharan Africa context. To this extent, we varied `c` within the range of 50 to 120 sexual acts per year based on data collected from serodiscordant couples across eastern and southern African sites. Previous research has shown considerable heterogeneity in the probability of HIV transmission per sexual contact, largely due to factors associated with the viral load level, genital ulcer disease, stage of HIV progression, condom use, circumcision and use of ART. Following a systematic review of this topic by Boily et al., we selected values for `beta` within the range of 0.003–0.008. ... Here, we chose values for `v` within the range of 0.15–0.35, which are slightly conservative, but supported by population-based estimates from the sub-Saharan African context."

`c` varies from 50 to 120 per year   
`beta` varies from 0.003 to 0.008  
`prev`, which here is population prevalence of unsuppressed VL, varies from 0.15 to 0.35  
`epsilon` can be initially parameterized using the RV144 Thai Trial results: VE = 61% at 12 months, 31% at 42 months


## Model setup

Basic functions:  

`dS/dt = -lambda*S`   
`dI/dt = lambda*S`  

Sp = susceptible placebo  
Ip = infected placebo  
Sv = susceptible vaccinated  
Iv = infected vaccinated  

Svh = susceptible vaccinated high exposure  
Svm = susceptible vaccinated medium exposure  
SvL = susceptible vaccinated low exposure (zero in this instance)  
Ivh = infected vaccinated high exposure  
Ivm = infected vaccinated medium exposure  
Ivl = infected vaccinated low exposure (zero in this instance)  

```{r parameters}

beta <- 0.004   #transmission rate (per contact)
c <- 90/365    #contact rate (contacts per day)
prev <- 0.10   #needs some more consideration
lambda <- beta*c*prev
epsilon <- 0.30 #per contact vaccine efficacy
risk <- 10.0 #risk multiplier
```
  
  
```{r ODE model}

si_ode <- function(times, init, param){
  with(as.list(c(init, param)), {
    
    # Flows
    # the number of people moving from S to I at each time step
    #Susceptible, Infected, placebo
    SIp.flow <- lambda*Sp
    SIv.flow <- lambda*(1-epsilon)*Sv
    
    #Susceptible, Infected, placebo, high, medium, low
    SIph.flow <- risk*lambda*Sph
    SIpm.flow <- lambda*Spl
    SIpl.flow <- 0*lambda*Spl  #0 to give this group zero exposures
    
    #Susceptible, Infected, vaccine, high, medium, low
    SIvh.flow <- risk*lambda*(1-epsilon)*Svh
    SIvm.flow <- lambda*(1-epsilon)*Spl
    SIvl.flow <- 0*lambda*(1-epsilon)*Svl  #0 to give this group zero exposures
    
    # ODEs
    # placebo; homogeneous risk
    dSp <- -SIp.flow
    dIp <- SIp.flow  #lambda*Sp
    
    # vaccine; homogeneous risk
    dSv <- -SIv.flow
    dIv <- SIv.flow  #lambda*epsilon*Sv

    # placebo; heterogeneous risk
    dSph <- -SIph.flow
    dIph <- SIph.flow  #risk*lambda*Sph
    dSpm <- -SIpm.flow
    dIpm <- SIpm.flow  #lambda*Spm
    dSpl <- -SIpl.flow
    dIpl <- SIpl.flow  #0*lambda*Spl
    
    # vaccine; heterogeneous risk
    dSvh <- -SIvh.flow
    dIvh <- SIvh.flow  #risk*lambda*(1-epsilon)*Svh
    dSvm <- -SIvm.flow
    dIvm <- SIvm.flow  #lambda*Svm
    dSvl <- -SIvl.flow
    dIvl <- SIvl.flow  #0*lambda*(1-epsilon)*Svl

    #Output
    list(c(dSp,dIp,
           dSv,dIv,
           dSph,dIph,
           dSpm,dIpm,
           dSpl,dIpl,
           dSvh,dIvh,
           dSvm,dIvm,
           dSvl,dIvl,
           SIp.flow,SIv.flow,
           SIph.flow,SIpm.flow,SIpl.flow,
           SIvh.flow,SIvm.flow,SIvl.flow))
  })
}
```
  
  
```{r EpiModel}

param <- param.dcm(lambda = lambda, epsilon = epsilon, risk = risk)
init <- init.dcm(Sp = 5000, Ip = 0,
                 Sv = 5000, Iv = 0,
                 Sph = 500, Iph = 0,    #placebo, high risk
                 Spm = 4000, Ipm = 0,   #placebo, medium risk
                 Spl = 500, Ipl = 0,    #placebo, low risk
                 Svh = 500, Ivh = 0,    #vaccine
                 Svm = 4000, Ivm = 0,   #vaccine
                 Svl = 500, Ivl = 0,    #vaccine
                 SIp.flow = 0, SIv.flow = 0, 
                 SIph.flow = 0, SIpm.flow = 0, SIpl.flow = 0,
                 SIvh.flow = 0, SIvm.flow = 0, SIvl.flow = 0)

control <- control.dcm(nsteps = 365*3, new.mod = si_ode)

mod <- dcm(param, init, control)
mod
```


```{r data}
mod <- mutate_epi(mod, total.Svh.Svm.Svl = Svh + Svm + Svl) #all susceptible in heterogeneous risk vaccine pop
mod <- mutate_epi(mod, total.Sph.Spm.Spl = Sph + Spm + Spl) #all susceptible in heterogeneous risk placebo pop

mod <- mutate_epi(mod, total.Ivh.Ivm.Ivl = Ivh + Ivm + Ivl) #all infected in heterogeneous risk vaccine pop
mod <- mutate_epi(mod, total.Iph.Ipm.Ipl = Iph + Ipm + Ipl) #all infected in heterogeneous risk placebo pop

mod <- mutate_epi(mod, total.SIvh.SIvm.SIvl.flow = SIvh.flow + SIvm.flow + SIvl.flow) #all infections per day in heterogeneous risk vaccine pop
mod <- mutate_epi(mod, total.SIph.SIpm.SIpl.flow = SIph.flow + SIpm.flow + SIpl.flow) #all infections in heterogeneous risk placebo pop

#Incidence estimate, per 100 person years

#Instantaneous incidence
mod <- mutate_epi(mod, rate.Vaccine = (SIv.flow/Sv)*365*100)
mod <- mutate_epi(mod, rate.Placebo = (SIp.flow/Sp)*365*100)
mod <- mutate_epi(mod, rate.Vaccine.het = (total.SIvh.SIvm.SIvl.flow/total.Svh.Svm.Svl)*365*100)
mod <- mutate_epi(mod, rate.Placebo.het = (total.SIph.SIpm.SIpl.flow/total.Sph.Spm.Spl)*365*100)

#Cumulative incidence
mod <- mutate_epi(mod, cumul.Sv = cumsum(Sv))
mod <- mutate_epi(mod, cumul.Sp = cumsum(Sp))
mod <- mutate_epi(mod, cumul.Svh.Svm.Svl = cumsum(total.Svh.Svm.Svl))
mod <- mutate_epi(mod, cumul.Sph.Spm.Spl = cumsum(total.Sph.Spm.Spl))

mod <- mutate_epi(mod, cumul.rate.Vaccine = (Iv/cumul.Sv)*365*100)
mod <- mutate_epi(mod, cumul.rate.Placebo = (Ip/cumul.Sp)*365*100)
mod <- mutate_epi(mod, cumul.rate.Vaccine.het = (total.Ivh.Ivm.Ivl/cumul.Svh.Svm.Svl)*365*100)
mod <- mutate_epi(mod, cumul.rate.Placebo.het = (total.Iph.Ipm.Ipl/cumul.Sph.Spm.Spl)*365*100)


#VE <- 1 - Relative Risk; this is VE for instantaneous incidence
mod <- mutate_epi(mod, VE1.inst = 1 - rate.Vaccine/rate.Placebo)
mod <- mutate_epi(mod, VE2.inst = 1 - rate.Vaccine.het/rate.Placebo.het)

#VE <- 1 - Relative Risk; this is VE from cumulative incidence
mod <- mutate_epi(mod, VE1.cumul = 1 - cumul.rate.Vaccine/cumul.rate.Placebo)
mod <- mutate_epi(mod, VE2.cumul = 1 - cumul.rate.Vaccine.het/cumul.rate.Placebo.het)

#VE again
#mod <- mutate_epi(mod, Cumul.inc = 1 - exp(-rate.Vaccine * control$timesteps))

mod
```

   
```{r plots}

par(mar = c(3,3,2,1), mgp = c(2,1,0))
plot(mod, y = c("Iv", "Ip", "total.Ivh.Ivm.Ivl", "total.Iph.Ipm.Ipl"), 
     alpha = 0.8, 
     main = "Cumulative infections",
     legend = "full")
plot(mod, y = c("SIv.flow", "SIp.flow", "SIvh.flow", "SIvm.flow", "SIvl.flow", "SIph.flow", "SIpm.flow", "SIpl.flow"), 
     alpha = 0.8, 
     main = "Daily infections, w/ all risk groups shown",
     legend = "full")
plot(mod, y = c("SIv.flow", "SIp.flow", "total.SIvh.SIvm.SIvl.flow", "total.SIph.SIpm.SIpl.flow"),
     alpha = 0.8, 
     main = "Daily infections, total mixed risk group",
     legend = "full")
plot(mod, y=c("rate.Vaccine", "rate.Placebo", "rate.Vaccine.het", "rate.Placebo.het"),
     alpha = 0.8,
     #ylim = c(0, 0.1),
     main = "Instantaneous incidence rate",
     legend = "full")
plot(mod, y=c("cumul.rate.Vaccine", "cumul.rate.Placebo", "cumul.rate.Vaccine.het", "cumul.rate.Placebo.het"),
     alpha = 0.8,
     #ylim = c(0, 0.1),
     main = "Cumulative incidence rate",
     legend = "full")
plot(mod, y=c("VE1.inst", "VE2.inst", "VE1.cumul", "VE2.cumul"),
     alpha = 0.8,
     main = "Vaccine efficacy",
     legend = "full")

# Manual plot legend example
plot(mod, y=c("VE1.inst", "VE2.inst", "VE1.cumul", "VE2.cumul"),
     alpha = 0.8,
     main = "Vaccine efficacy",
     legend = FALSE, 
     col = 1:4)
legend("topright", legend = c("Inst VE 1", "Inst VE 2", "Cuml VE 1", "Cuml VE 2"),
       col = 1:4, lwd = 2, cex = 0.9, bty = "n")

```


